---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# spflow

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/spflow)](https://CRAN.R-project.org/package=spflow)
[![R-CMD-check](https://github.com/LukeCe/spflow/workflows/R-CMD-check/badge.svg)](https://github.com/LukeCe/spflow/actions)
[![test-coverage](https://codecov.io/gh/LukeCe/spflow/branch/master/graph/badge.svg)](https://github.com/LukeCe/spflow/actions)
[![status](https://tinyverse.netlify.com/badge/spflow)](https://CRAN.R-project.org/package=spflow)
[![Downloads](https://cranlogs.r-pkg.org/badges/spflow?color=brightgreen)](https://www.r-pkg.org/pkg/spflow)
<!-- badges: end -->


The [**spflow**](https://lukece.github.io/spflow/) package allows to estimate spatial econometric interaction models.
It is designed to exploit the relational structure of flow data, reducing the computational burden and memory requirements.

## Installation

You can install the released version of spflow from [CRAN](https://CRAN.R-project.org/package=spflow) with:

``` r
install.packages("spflow")
```

And the  development version from [GitHub](https://github.com/LukeCe/spflow/) with:

``` r
# install.packages("devtools")
devtools::install_github("LukeCe/spflow@non-cartesian-flows")
```

## Example

The package provides a new `spflow_network_multi-class` that combines information on the origins, the destinations, and the origin-destination pairs.
Once this object is created, we can estimate an interaction model with the `spflow()` function. 
By default the model includes three autoregression parameters: rho_o, rho_d, rho_w.
These parameters are related to origin-, destination-, and origin-to-destination-dependence.

Through the formula interface we specify which variables are used as origin `O_()`, destination `D_()`, intra-regional `I_()` and OD-pair `P_()` characteristics.

For more detailed examples have a look at the [package vignette](https://lukece.github.io/spflow/articles/paris_commute_flows.html).


```{r, message=FALSE}
library("spflow")
data("multi_net_usa_ge")

spflow(y9 ~ O_(X) + D_(X) + I(X) + P_(DISTANCE), multi_net_usa_ge)
```


## License

[GPL 3](https://www.r-project.org/Licenses/GPL-3)
