---
title: "Home-to-work commuting flows within the municipalities around Paris"
output: 
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{Home-to-work commuting flows within the municipalities around Paris}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(max.print = 200)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library("spflow")
library("sf")
library("spdep")
```

# Introduction

This article illustrates the use of the `spflow` package for spatial interaction modelling, using the example of home-to-work commuting flows within 71 municipalities that are no further than 10km away from the center of Paris.
The examples employs data to describe the municipalities their neighborhood structure additional data on the size of the commuting flow and the distance between all possible origin-destination pairs is also available.


```{r}
data("paris10km_municipalities")
data("paris10km_neighborhood")
data("paris10km_commuteflows")
```

# Data presentation

Each municipality has a unique identification and additional variables describing it.

```{r}
drop_area <- names(paris10km_municipalities) != "AREA"
plot(paris10km_municipalities[drop_area])
```


Additionally there are three different neighborhood matrices that can be used to describe the connectivity between the municipalities.

```{r}
par(mfrow = c(1, 3), mar = c(0,0,1,0))
mid_points <- suppressWarnings({
    st_point_on_surface(st_geometry(paris10km_municipalities))})

paris10km_nb <- list(
  "by_contiguity" = spdep::poly2nb(paris10km_municipalities),
  "by_distance" = spdep::dnearneigh(mid_points,d1 = 0, d2 = 5),
  "by_knn" = spdep::knn2nb(knearneigh(mid_points,3))
)

plot(st_geometry(paris10km_municipalities))
plot(paris10km_nb$by_contiguity, mid_points, add = T, col = rgb(0,0,0,alpha=0.5))
title("Contiguity") 

plot(st_geometry(paris10km_municipalities))
plot(paris10km_nb$by_distance,mid_points, add = T, col = rgb(0,0,0,alpha=0.5)) 
title("Distance") 

plot(st_geometry(paris10km_municipalities))
plot(paris10km_nb$by_knn, mid_points, add = T, col = rgb(0,0,0,alpha=0.5))
title("3 Nearest Neighbors") 
```

Finally there is data on the size of the commuting flows between pairs of origins and destinations

```{r}
paris10km_commuteflows 
```


# Modeling Spatial interactions

The `spflow` package build on the idea that any type of flow corresponds to pair-wise interactions between the nodes of an origin network with the nodes of a destination network.

In our example the origin and destination networks are the same since each of the municipalities is at the same time the origin and the destination of home-to-work commuting flows.


For efficient estimation of the model the `spflow` package employs a moment based estimation method.
This avoids duplication which arises due to the fact that every municipality is at the origin and the destination of many flows.
Using moment based estimator reduces the size of the data for most parts of the estimation to $n$ while the traditional method will create information for all origin-destination pairs $N = n^2$. 

## The network objects 

To describe the nodes of the network we then use the `sp_network_nodes` object which combines attributes of the nodes with the chosen network structure.

```{r}
paris10km_net <- 
  sp_network_nodes(
    network_id = "paris10km",
    node_neighborhood = nb2mat(paris10km_nb$by_contiguity),
    node_data = st_drop_geometry(paris10km_municipalities),
    node_key_column = "ID_MUN")

paris10km_net
```

The `sp_network_pairs` object contains all information to describe the pairs of the origin and destination networks.

```{r}
paris10km_net_pairs <- sp_network_pair(
  orig_net_id = "paris10km",
  dest_net_id = "paris10km",
  pair_data = paris10km_commuteflows,
  orig_key_column = "ID_ORIG",
  dest_key_column = "ID_DEST")

paris10km_net_pairs
```

The `sp_multi_network` object is than used to gather and ensure consistency of information between the information on the single nodes and the information on the node pairs.
It contains all information on origins destinations and origin-destination pairs and no further data needs to be passed to the main modeling function.

```{r}
paris10km_multi_net <- sp_multi_network(paris10km_net,paris10km_net_pairs)
paris10km_multi_net

```

## Estimating with default values

Estimation with default settings requires two arguments: a `sp_multi_inetwork` which contains all the required data and a `flow_formula`.
The `flow_formula` specifies the model we want to estimate.
In this example the dependent variable is a transformation of the commuting flows and we use the dot-shortcut to indicate that all available variables should be included in the model.
lead by default to the most comprehensive spatial interaction model that the package allows to estimate.

```{r}
sim_fit_default <- spflow(
  flow_formula = log(1 + COMMUTE_FLOW) ~ . + G_(log( 1 + DISTANCE)),
  sp_multi_network = paris10km_multi_net)


sim_fit_default
```

## Adjusting the formula

In the previous example we used the default estimation leads to a spatial interaction model which uses all the available variables.


```{r}
formula_1 <- 
  log(1 + COMMUTE_FLOW) ~ 
  AREA + MED_INCOME + log(POPULATION + 1 ) + log(NB_COMPANY + 1) + 
  G_( log(DISTANCE + 1) )

control_1 <- spflow_control(use_intra = FALSE)

sim_fit_1 <- 
  spflow(flow_formula = formula_1,
         sp_multi_network = paris10km_multi_net,
         flow_control = control_1)

sim_fit_1
```



