---
title: "Home-to-work commuting flows within the municipalities around Paris"
output: 
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{Home-to-work commuting flows within the municipalities around Paris}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(max.print = 200)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library("spflow")
library("sf")
library("spdep")
```

# Introduction

This article illustrates the use of the **spflow** package for modeling spatial interactions using the example of home-to-work commuting flows within 71 municipalities located no more than 10km from the center of Paris.
The examples use data describing the municipalities and their neighborhood structure. Additional data on the size of the commuter flow and the distance between all possible origin-destination pairs are also available.


```{r}
data("paris10km_municipalities")
data("paris10km_neighborhood")
data("paris10km_commuteflows")
```

# Data presentation

Each municipality has a unique identification and additional variables describing it.

```{r}
drop_area <- names(paris10km_municipalities) != "AREA"
plot(paris10km_municipalities[drop_area])
```


Additionally, there are three different neighborhood matrices that can be used to describe the connectivity between the municipalities.

```{r}
par(mfrow = c(1, 3), mar = c(0,0,1,0))
mid_points <- suppressWarnings({
    st_point_on_surface(st_geometry(paris10km_municipalities))})

paris10km_nb <- list(
  "by_contiguity" = spdep::poly2nb(paris10km_municipalities),
  "by_distance" = spdep::dnearneigh(mid_points,d1 = 0, d2 = 5),
  "by_knn" = spdep::knn2nb(knearneigh(mid_points,3))
)

plot(st_geometry(paris10km_municipalities))
plot(paris10km_nb$by_contiguity, mid_points, add = T, col = rgb(0,0,0,alpha=0.5))
title("Contiguity") 

plot(st_geometry(paris10km_municipalities))
plot(paris10km_nb$by_distance,mid_points, add = T, col = rgb(0,0,0,alpha=0.5)) 
title("Distance") 

plot(st_geometry(paris10km_municipalities))
plot(paris10km_nb$by_knn, mid_points, add = T, col = rgb(0,0,0,alpha=0.5))
title("3 Nearest Neighbors") 
```

Finally, there is data on the size of the commuting flows between pairs of origins and destinations.

```{r}
paris10km_commuteflows 
```


# Modeling Spatial interactions

The **spflow** package builds on the idea that each type of flow corresponds to a pairwise interaction between the nodes of an origin network with the nodes of a destination network.

In our example, the origin and destination networks are the same because every municipality is both an origin and destination of a flow.

To estimate the model efficiently, the **spflow** package uses moment-based estimation methods, that exploit a matrix formulation of the model.
This avoids duplication caused by each community being both the origin and destination of many flows.


## The network objects 

To describe the nodes of the network the package provides `sp_network_nodes-class` that combines attributes of the nodes with the chosen network structure.

```{r}
paris10km_net <- 
  sp_network_nodes(
    network_id = "paris10km",
    node_neighborhood = nb2mat(paris10km_nb$by_contiguity),
    node_data = st_drop_geometry(paris10km_municipalities),
    node_key_column = "ID_MUN")

paris10km_net
```

The `sp_network_pair-class` contains all information to describe the pairs of the origin and destination networks.

```{r}
paris10km_net_pairs <- sp_network_pair(
  orig_net_id = "paris10km",
  dest_net_id = "paris10km",
  pair_data = paris10km_commuteflows,
  orig_key_column = "ID_ORIG",
  dest_key_column = "ID_DEST")

paris10km_net_pairs
```

The `sp_multi_network-class` is than combines information on the nodes and the node-pairs and ensures also ensures that both data sources are consistent.
For example, if some of the origins in the `sp_network_pair-class` are not identified with the nodes in the `sp_network_nodes-class` an error will be shown.

```{r}
paris10km_multi_net <- sp_multi_network(paris10km_net,paris10km_net_pairs)
paris10km_multi_net
```

## Estimation

The core function of the package is `spflow()`, which provides an interface to four different estimators of the spatial econometric interaction model.

### Estimating with default values

Estimation with default settings requires two arguments: an `sp_multi_network-class` and a `flow_formula`.
The `flow_formula` specifies the model we want to estimate.
In this example, the dependent variable is a transformation of commuting flows and we use the dot shortcut to indicate that all available variables should be included in the model.
Using the defaults leads to the most comprehensive spatial interaction model, which includes spatial lags of the dependent variable, the exogenous variables and additional attributes for intra-regional observations.

```{r}
results_default <- spflow(
  flow_formula = log(1 + COMMUTE_FLOW) ~ . + G_(log( 1 + DISTANCE)),
  sp_multi_network = paris10km_multi_net)

results_default
```

### Adjusting the formula

We can adjust how the exogenous variables are to be used by wrapping them into the `D_()`, `O_()`, `I_()` and `G_()` functions.
The variables in `G_()` are used as OD pair features and those in `D_()`, `O_()` and `I_()` are used as destination, origin and intra-regional features.
We can take advantage of the formula interface to specify transformations and expand factor variables to dummies.


```{r}
clog <- function(x) {
  log_x <- log(x)
  log_x - mean(log_x)
}

flow_formula  <- 
  log(COMMUTE_FLOW + 1) ~
  D_(log(NB_COMPANY) + clog(MED_INCOME)) +
  O_(log(POPULATION) + clog(MED_INCOME)) +
  I_(log(POPULATION)) +
  G_(log(DISTANCE + 1))

results_mle  <- spflow(
  flow_formula,
  paris10km_multi_net)
results_mle
```


### Adjustment through `spflow_control()`

More fine-grained adjustments are possible via the `spflow_control()` argument.
Here we change the estimation method and the way we want to model the spatial autoregression in the flows.
To use spatial lags only for certain variables, we need to specify them as a second formula. 


```{r}
sdm_formula <- ~
  O_(log(POPULATION) + clog(MED_INCOME)) +
  D_(log(NB_COMPANY) + clog(MED_INCOME))

cntrl <- spflow_control(
  estimation_method = "mcmc",
  sdm_variables = sdm_formula,
  model = "model_7") # restricts \rho_w = 0

results_mcmc  <- spflow(
  flow_formula,
  paris10km_multi_net,
  flow_control = cntrl)

results_mcmc
```

